---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = '#>',
  fig.path = 'man/figures/README-',
  out.width = '100%'
)
```

# PAGFL

<!-- badges: start -->

[![CRAN_Version_Badge](http://www.r-pkg.org/badges/version/PAGFL)](https://cran.r-project.org/package=PAGFL) [![CRAN_Downloads_Badge](https://cranlogs.r-pkg.org/badges/grand-total/PAGFL)](https://cran.r-project.org/package=PAGFL) [![License_GPLv3_Badge](https://img.shields.io/badge/License-GPLv3-yellow.svg)](https://www.gnu.org/licenses/gpl-3.0.html) [![R-CMD-check](https://github.com/Paul-Haimerl/PAGFL/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Paul-Haimerl/PAGFL/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

In panel data analysis, unobservable group structures are a common challenge. Disregarding group-level heterogeneity by assuming an entirely homogeneous panel can introduce bias. Conversely, estimating individual coefficients for each cross-sectional unit is inefficient and may lead to high uncertainty.

Mehrabani ([2023](https://doi.org/10.1016/j.jeconom.2022.12.002)) introduces the pairwise adaptive group fused Lasso (*PAGFL*), a fast methodology to identify latent group structures and estimate group-specific coefficients simultaneously.

The `PAGFL` package makes this powerful procedure easy to use.

## Installation

You can install the development version of `PAGFL` (1.1.0) from [GitHub](https://github.com/) with:

```{r message=FALSE, warning=FALSE}
# install.packages('devtools')
devtools::install_github('Paul-Haimerl/PAGFL')
library(PAGFL)
```

The stable version (1.0.1) is available on CRAN:

```         
install.packages("PAGFL")
```

## Data

The `PAGFL` packages includes a function that automatically simulates a panel with a group structure:

```{r}
# Simulate a simple panel with three distinct groups and two exogenous explanatory variables
set.seed(1)
sim <- sim_DGP(N = 50, n_periods = 150, p = 2, n_groups = 3)
y <- sim$y
X <- sim$X
```

$$y_{it} = \beta_i^\prime x_{it} + \eta_i + u_{it}, \quad i = 1, \dots, N, \quad t = 1, \dots, T,$$ where $y_{it}$ is a scalar dependent variable, $x_{it}$ a $p \times 1$ vector of explanatory variables, and $\eta_i$ reflects a fixed effect. The slope coefficients are subject to the group structure

$$\beta_{i} = \sum_{k = 1}^K \alpha_k \boldsymbol{1} \{i \in G_k \},$$ with $\cup_{k = 1}^K G_k = \{1, \dots, N \}$, and $G_k \cap G_j = \emptyset$ as well as $|| \alpha_k \neq \alpha_j ||$ for any $k \neq j$, $k,j = 1, \dots, K$ (see Mehrabani [2023](https://doi.org/10.1016/j.jeconom.2022.12.002), sec. 2).

`sim_DGP` also nests, among other, all DGPs employed in the simulation study of Mehrabani ([2023](https://doi.org/10.1016/j.jeconom.2022.12.002), sec. 6). I refer to the documentation of `sim_DGP` or Mehrabani ([2023](https://doi.org/10.1016/j.jeconom.2022.12.002), sec. 6) for more details.

## Applying PAGFL

To execute the PAGFL procedure, simply pass the dependent and independent variables, the number of time periods, and a penalization parameter $\lambda$.

```{r}
estim <- pagfl(y = y, X = X, n_periods = 150, lambda = 10)
print(estim)
```

`pagfl` returns a list holding

1.  the value of the IC (see Mehrabani [2023](https://doi.org/10.1016/j.jeconom.2022.12.002), sec. 3.4)
2.  the $\lambda$ parameter
3.  the $\widehat{K} \times p$ post-Lasso coefficient matrix $\hat{\boldsymbol{\alpha}}^p$
4.  the estimated number of groups $\widehat{K}$
5.  the estimated group structure
6.  the number of executed *ADMM* algorithm iterations
7.  a logical indicator if convergence was achieved

Selecting a $\lambda$ value a priori can be tricky. For instance, it seems like `lambda = 10` is too high since the number of groups $K$ is underestimated. As a consequence, we suggest iterating over a comprehensive range of candidate values. To specify a suitable grid, create a logarithmic sequence ranging from 0 to a penalty parameter that induces an entirely homogeneous model (i.e., $\widehat{K} = 1$). The resulting $\lambda$ grid vector can be passed in place of any specific value, and a BIC IC selects the best-fitting parameter.

Moreover, if the explanatory variables in `X` are named, those names also appear in the output.

```{r}
colnames(X) <- c("a", "b")
lambda_set <- exp(log(10) * seq(log10(1e-4), log10(10), length.out = 10))
estim_set <- pagfl(y = y, X = X, n_periods = 150, lambda = lambda_set)
print(estim_set)
```

When, as above, the specific estimation method is left unspecified, `pagfl` defaults to penalized Least Squares (*PLS*) `method = 'PLS'` (Mehrabani, [2023](https://doi.org/10.1016/j.jeconom.2022.12.002), sec. 2.2). *PLS* is very efficient but requires weakly exogenous regressors. However, even endogenous predictors can be accounted for by employing a penalized Generalized Method of Moments (*PGMM*) routine in combination with exogenous instruments $\boldsymbol{Z}$.

Specify a slightly more elaborate endogenous and dynamic panel data set and apply *PGMM*. When encountering a dynamic panel data set, we recommend using a Jackknife bias correction, as proposed by Dhaene and Jochmans ([2015](https://doi.org/10.1093/restud/rdv007)).

```{r}
# Generate a panel where the predictors X correlate with the cross-sectional innovation, 
# but can be instrumented with q = 3 variables in Z. Furthermore, include GARCH(1,1) 
# innovations, an AR lag of the dependent variable, and specific group sizes
sim_endo <- sim_DGP(N = 50, n_periods = 150, p = 2, n_groups = 3, group_proportions = c(0.2, 0.2, 0.6), 
error_spec = 'GARCH', q = 3)
y_endo <- sim_endo$y
X_endo <- sim_endo$X
Z <- sim_endo$Z

# Note that the method PGMM and the instrument matrix Z needs to be supplied
estim_endo <- pagfl(y = y_endo, X = X_endo, n_periods = 150, lambda = 0.05, method = 'PGMM', Z = Z, 
bias_correc = TRUE, max_iter = 8e3)
print(estim_endo)
```

Furthermore, `pagfl` lets you select a minimum group size, adjust the efficiency vs. accuracy trade-off of the iterative estimation algorithm, and modify a list of further settings. Visit the documentation `?pagfl()` for more information.

## The Time-varying PAGFL

The development version of the package also includes the functions `sim_dyn_DGP`and `tv_pagfl`, which generate and estimate a grouped panel data models with the time-varying coefficients $\beta_{it} = \beta_i \left( \frac{t}{T} \right)$. Just like in the static case, the functional coefficients admit to a group structure $\beta_{it} = \sum_{k = 1}^K \alpha_k \left( \frac{t}{T} \right) 1 \{i \in G_k \}$. Following Su et al. ([2019](https://doi.org/10.1080/07350015.2017.1340299)), the time-varying coefficients are estimated using polynomial B-spline functions employing a penalized sieve estimation (*PSE*).

```{r}
# Simulate a time-varying panel with a trend and a group pattern
N <- 50
n_periods <- 50
sim_dyn <- sim_tv_DGP(N = N, n_periods = n_periods, DGP = 1)
y <- sim_dyn$y
X <- sim_dyn$X

dyn_estim <- tv_pagfl(y = y, X = X, n_periods = n_periods, lambda = 6)
```

In contrast to the time-constant `pagfl`, `tv_pagfl` does not return a post-Lasso coefficient matrix but a $T \times p \times \widehat{K}$ array.

In empirical applications, it is commonplace to encounter unbalanced panel data sets. In such instances, time-varying coefficient functions can be estimated nonetheless. The nonparametric spline functions simply smooth over the time periods with only a limited number or even no observations. However, it is required to provide explicit indicator variables that declare the cross-sectional individual and time period each observation belongs to.

Lets delete a couple of observations, add indicator variables, and run `tv_pagfl` again.

```{r}
# Draw some observations to be omitted
delete_index_y <- as.logical(rbinom(n = N * n_periods, prob = 0.9, size = 1))
delete_index_X <- as.logical(rbinom(n = N * n_periods, prob = 0.9, size = 1))
# Construct cross-sectional and time indicator variables
i_index <- rep(1:N, each = n_periods)
t_index <- rep(1:n_periods, N)
y <- cbind(y, i_index = i_index, t_index = t_index)
X <- cbind(X, i_index = i_index, t_index = t_index)
# Delte some observations
y <- y[delete_index_y,]
X <- X[delete_index_X,]
# Apply the time-varying PAGFL to an unbalanced panel
dyn_estim_unbalanced <- tv_pagfl(y = y, X = X, index = c("i_index", "t_index"), lambda = 6)
```



Furthermore, `tv_pagfl` lets you specify a lot more optionalities than shown here. For example, it is possible to adjust the degree of the spline functions, the number of interior knots in the spline system, or estimate a panel data model with a mix of time-varying and time-constant coefficients. See `?tv_pagfl()` for details.

## References

-   Dhaene, G., & Jochmans, K. (2015). Split-panel jackknife estimation of fixed-effect models. *The Review of Economic Studies*, 82(3), 991-1030. DOI: [10.1093/restud/rdv007](https://doi.org/10.1093/restud/rdv007)

-   Mehrabani, A. (2023). Estimation and identification of latent group structures in panel data. *Journal of Econometrics*, 235(2), 1464-1482. DOI: [10.1016/j.jeconom.2022.12.002](https://doi.org/10.1016/j.jeconom.2022.12.002)

-   Schumaker, L. (2007). Spline functions: basic theory. *Cambridge university press*. DOI: [10.1017/CBO9780511618994](https://doi.org/10.1017/CBO9780511618994)

-   Su, L., Wang, X., & Jin, S. (2019). Sieve estimation of time-varying panel data models with latent structures. *Journal of Business & Economic Statistics*, 37(2), 334-349. DOI: [10.1080/07350015.2017.1340299](https://doi.org/10.1080/07350015.2017.1340299)
