---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PAGFL

<!-- badges: start -->
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![R-CMD-check](https://github.com/Paul-Haimerl/PAGFL/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Paul-Haimerl/PAGFL/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

In panel data analysis, the presence of unobservable group structures is a common challenge. Disregarding group-level heterogeneity by assuming a fully homogeneous panel can introduce bias. Conversely, estimating individual coefficients for each cross-sectional unit is inefficient and may lead to high uncertainty.

Mehrabani ([2023](https://doi.org/10.1016/j.jeconom.2022.12.002)) introduces the pairwise adaptive group fused Lasso (PAGFL), a fast methodology to simultaneously identify latent group structures and estimate group-specific coefficients.

The `PAGFL` package makes it easy to use this powerful procedure.

## Installation

You can install the development version of `PAGFL` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Paul-Haimerl/PAGFL")
```

After installation, attach the package as usual:

```{r}
library(PAGFL)
```


## Data

The `PAGFL` packages includes a function that automatically simulates a panel with a group structure:

```{r}
# Simulate a simple panel with three distinct groups and two exogenous explanatory variables
set.seed(2)
sim <- sim_DGP(N = 50, n_periods = 80, p = 2, n_groups = 3)
y <- sim$y
X <- sim$X
```

`sim_DGP` also nests, among other, all DGPs employed in the simulation study of Mehrabani ([2023](https://doi.org/10.1016/j.jeconom.2022.12.002), sec. 6). I refer to the documentation of `sim_DGP` or Mehrabani ([2023](https://doi.org/10.1016/j.jeconom.2022.12.002), sec. 6) for more details.

## Applying PAGFL

To execute the PAGFL procedure, simply pass the dependent and independent variables, number of time periods and a penalization parameter $\lambda$.

```{r}
estim <- PAGFL(y = y, X = X, n_periods = 80, lambda = 5)
print(estim)
```

Selecting a $\lambda$ value a priori can be tricky. Therefore, we suggest iterating over a comprehensive range of candidate values. To specify a suitable grid, create a logarithmic sequence ranging from 0 to a penalty parameter that induces a fully homogeneous model. The resulting $\lambda$ grid vector can be simply passed in place of any specific value and a BIC IC selects the best fitting parameter.

```{r}
lambda_set <- exp(log(10) * seq(log10(1e-4), log10(10), length.out = 10))
estim_set <- PAGFL(y = y, X = X, n_periods = 80, lambda = lambda_set)
print(estim_set)
```

When, as above, the specific estimation method is left unspecified, `PAGFL` defaults to penalized Least Squares (*PLS*) (Mehrabani, [2023](https://doi.org/10.1016/j.jeconom.2022.12.002), sec. 2.2). *PLS* is very efficient, but requires weakly exogenous regressors. However, even endogenous predictors can be accounted for by employing a penalized Generalized Method of Moments (*PGMM*) routine in combination with exogenous instruments $Z$.

Lets specify a slightly more elaborate endogenous panel data set and apply *PGMM*.

```{r}
# Generate a panel where the predictors X correlate with the cross-sectional innovation, but can be instrumented with q = 3 variables in Z
sim_endo <- sim_DGP(N = 50, n_periods = 80, p = 2, n_groups = 3, q = 3)
y_endo <- sim_endo$y
X_endo <- sim_endo$X
Z <- sim_endo$Z

# Note that a method PGMM and the instrument matrix Z needs to be supplied
estim_endo <- PAGFL(y = y, X = X, n_periods = 80, lambda = 0.05, method = 'PGMM', Z = Z)
print(estim_endo)
```

Furthermore, `PAGFL` lets you select a minimum group size, toggle a Split-panel Jackknife bias correction and adjust a list of additional setting. Visit the documentation of `PAGFL` for more information.

## References

Mehrabani, A. (2023). Estimation and identification of latent group structures in panel data. *Journal of Econometrics*, 235(2), 1464-1482. DOI: [10.1016/j.jeconom.2022.12.002](https://doi.org/10.1016/j.jeconom.2022.12.002)
